Exploiting the Graph: Increasing Engagement in Collaborative Interactive Evolution

### 2017-01-23 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
More

We present in this section the results of a comparison between two versions
of the same IEC application. The goal of the experiment is to compare two 
gamification techniques to increase volunteer participation. 
The main difference between both versions is how the fitness of individuals
is calculated and the information presented to users. 
\subsection{Gamification technique}
\label{sec:gamification}
The gamification technique
consists in giving more importance to the preference of users with more
participations. A feature found in video games is the concept of a 
score and experience levels. Levels depend on the score and certain features
of the game are only available to gamers when they reach a giving level.

The score achieved by users depends on the actions he does in the system as
each time a user does one of these actions their score is incremented by one.
The actions considered in this application are: 
\begin{itemize}
\item Start a session.
\item Rate an individual.
\item Create a collection.
\item Save an individual of the wall to a collection.
\item Save an individual from a friend's collection.
\item Explore collections of other friends.
\end{itemize}
Only the "Start a session" and "Explore collections of other friends". 
Are not found in the graph model, these are stored in the PostgreSQL relational
database system.

Two scores are used to determine the weight of a user's preference:
\begin{itemize}
\item Experience. This value depends on the score and is a value 
between 0 and 100.

@@

The three versions of the EvoDrawing web application were deployed to the
Heroku cloud service, using the same deployment configuration, using the
Heroku Free option with a 512 MB of RAM, adding only the GrapheneDB plug-in for the GG experiment.
Table \ref{tab:params} shows the parameters used for the evolutionary algorithm.

@@

% Anuncements
In order to start each one of the experiments was promoted through social networks particularly on the Facebook
and Twitter platforms. The link used in the call for participation  was shortened URL, in order to obtain additional
analytic data of the visitors.  In table \ref{tab:urls} the URL for each deployment in the Heroku platform is shown,
along with the short URL and the analytics link. In the same table a link to the GitHub application repository
used to deploy to Heroku is also listed.

@@

Only data for the first week of deployment was considered for the experiments, and they where conducted 
between January and May of 2016. Each deployment was tested first with a few beta testers, before launch
in order to test design decisions.

% Time
\subsection{Results}
\label{sec:results}

>>>>

We present in this section the results of a comparison between three versions
of the same IEC application. The goal of the experiment is to compare two 
gamification techniques to increase volunteer participation against a third 
without any thechnique applied.

The main difference between both versions employing gamification is how the fitness of individuals
is calculated and the information presented to users.

\subsection{Gamification techniques}
\label{sec:gamification}

Gamification is defined by Deterding et al. as
``the use of game design elements in non-game contexts'' \cite{deterding2011game}.
The gamification element employed in this work is rewarding mechanism 
\cite{dubois2013understanding}, in general they consisting of a reputation systems with score points, 
levels and leader boards. Points are awarded to users in response of
the accomplishment of certain activities that need to be encouraged. Levels depend
on the score and certain features of the game are only available to gamers when 
they reach a giving level.

The rewarding mechanism as it is applied in these versions 
consists in giving more importance to the preference of those users having a higher reputation
as given by the score points and experience levels.

The score achieved by users depends on the actions he does in the system as
each time a user does one of these actions their score is incremented by one:
\begin{itemize}
\item Start a session.
\item Rate an individual.
\item Create a collection.
\item Save an individual of the wall to a collection.
\item Save an individual from a friend's collection.
\item Explore collections of other friends.
\end{itemize}

Only the "Start a session" and "Explore collections of other friends". 
Are not found in the graph model, these are stored in the PostgreSQL relational
database system.

Two scores are used to determine the weight of a user's preference:
\begin{itemize}
\item Experience. This value depends on the score and is a value 
between 0 and 100. 
% More details needed? Is done by a Fuzzy Inference System but I think
% the description would take to much space.

@@

The three versions of the EvoDrawing web application were deployed to the
Heroku cloud service with the same configuration: using the
Heroku Free option with a 512 MB of RAM, adding only the GrapheneDB plug-in for the GG experiment.
Table \ref{tab:params} shows the parameters used for the evolutionary algorithm. The graph database 
system used in the implementation was Neo4J, and it was allready available as a plug-in in Heroku.

@@

% Anuncements
In order to start each of the experiments was promoted through social networks
particularly on the Facebook and Twitter platforms. The link used in the call for participation
was shortened URL, in order to obtain additional analytic data of the visitors.
In table \ref{tab:urls} the URL for each deployment in the Heroku platform is shown,
along with the short URL and the analytics link. In the same table a link to the GitHub 
application repository used to deploy to Heroku is also listed.

@@

Only data for the first week of deployment was considered for the experiments, and they where conducted 
between January and May of 2016.

% Time
\subsection{Results}
\label{sec:results}
Before launch, each deployment was first tried with a few beta testers. When applying the gammification
technique for the first time a problem was found in this beta stage. Some users were chating by 
giving a rating to an animation even before it was returned from the server, this was done by just
constantly pressing the mouse button. This is a common problem found in systmes appling leader boards.
By making the rewards for tasks achievements visible to other players or providing leader boards are ways of 
encouraging players to compete \cite{hickman2010total}. Because the  problematic consequences of competition, 
which can result in negative conduct, low cooperation and collaboration, or disadvantaging certain player
demographics such as women \cite{kumar2013gamification}, best-practice gamification designs try to
refrain from using this element. In order to measure the effect of the leader board in the system, the 
problem of cheating was mitigated by disabling the buttons for a certain amount of time. But the problem 
persists, competitive users will try to get in to the leader board by other means. 
% 
% Future work could compare the use of gamification with out leaderboards
%

Using a graph database provided certain advantages when monitoring the experiments, for instance 
using the query language
